---
title: "Load and clean ASA24 data"
output: html_document
code_folding: hide
---

### What this script does:
1. Use Metadata 1 to filter out individuals.
2. Remove users that has only a small number of totals (days of record) - if you know which one to remove.  
3. Look for outliers in your totals by nutrient consumed on each day.

Name the path to DietR directory where input files are pulled.
```{r}
  main_wd <- "~/GitHub/DietR"
```

Load the necessary functions
```{r}
  source("lib/specify_data_dir.R")
  source("lib/load_clean_ASA24.R")
  source("lib/format.file.R")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_knit$set(root.dir = 'eg_data/VVKAJ')
```

### Load ASA24 Items data

Specify the directory where the data is.
```{r, eval = FALSE}
  SpecifyDataDirectory(directory.name= "eg_data/VVKAJ/")
```
<!-- [NOTE] SpecifyDataDirectory function is not executed by R markdown, but just showing  -->
<!--  the code so that this tutorial will have the same code as the R script.  -->

<!-- Define the path to items data in main_wd. -->
<!-- ```{r} -->
<!-- items_path <- paste("/eg_data/VVKAJ/VVKAJ_Items.csv", sep="") -->
<!-- print(items_path) -->
<!-- ``` -->

Load items data which is in a csv format.
```{r}
  items_raw <- read.csv("VVKAJ_Items.csv", sep = ",", header=T)
```

Show the first two lines of the items data.
```{r}
  head(items_raw, 2)
```

Save it as a .txt file.

This will be saved in GitHub/DietR/eg_data/VVKAJ directory.
```{r}
  write.table(items_raw, "VVKAJ_Items.txt", sep="\t", row.names=F)
```

Special characters such as "'", ",", "%" may interfere correct data loading; thus, we replace them with an underscore "_".

This function can only take .txt files as input. Specify column(s) to be processed in the "columns" argument and the output file name in the outfn argument; "_f" stands for "formatted".
```{r}
  format.file(filename = "VVKAJ_Items.txt",
              columns  = "Food_Description",
              outfn    = "VVKAJ_Items_f.txt")
```

Add SampleID with a desired prefix, and save it as a txt file. SampleIDs are IDs unique to each combination of users and day.
```{r}
  AddSampleIDtoItems(input.fn="VVKAJ_Items_f.txt", user.name="UserName", recall.no="RecallNo",
                     prefix="vvkaj.", out.fn="VVKAJ_Items_f_id.txt")
```

Load the formatted Items file.
```{r}
  items_f_id <- read.table("VVKAJ_Items_f_id.txt", sep="\t", header=T)
```

A combination of the specified prefix and sequential number should be added.
```{r, eval=FALSE}
  head(items_f_id)
```

<details>
  <summary>Click to expand output</summary>
```{r, echo=FALSE, eval=TRUE}
head(items_f_id)
```
</details>

<br>

Ensure your items file has the expected dimensions (number of rows x number of columns, shown as number of obs. and number of variables) in the environment window of R Studio, or you can also check the dimension of items_f by using the dim() function.
```{r}
  dim(items_f_id)
```

<br>

### <Optional> Use individuals_to_remove.txt to filter out users marked as Remove = yes.
Load your metadata that has information about which UserName(s) to remove.
```{r}
ind_to_rm <- read.table("individuals_to_remove.txt", sep="\t", header=T)
```

Take a look. Metadata for this purpose (ind_to_rm) has UserName and which one to be removed:
```{r}
  ind_to_rm
```

Show which has "yes" in the "Remove" column.
```{r}
  subset(ind_to_rm, Remove == "yes")

```

Remove the specified individuals.
The output will be saved as a text file with the specified name.
This assumes the usernames are in UserName column, and will print which user(s) will be removed.
```{r}
  RemoveRows(data=items_f_id, metadata.file= ind_to_rm,
             output.name= "VVKAJ_Items_f_id_s.txt")
```

Load the output for further processing.
```{r}
  items_f_id_s <- read.table("VVKAJ_Items_f_id_s.txt", header=T, sep="\t")
```

Show unique usernames in items_f_id_s and confirm "VVKAJ116" has been removed.
```{r}
  unique(items_f_id_s$UserName)
```

<br>

### <Optional> Merge individuals' metadata to items.   

ind_metadata has the participants' gender, age, height, weight, BMI, and Waist.Circumference, etc. If desired, this individual-specific information can be added to items data.
  
Load ind_metadata.txt.
```{r}
  ind_metadata <- read.table("ind_metadata.txt", sep="\t", header=T)
```

Look at what the metadata has.
```{r}
  head(ind_metadata)
```
  
Add this metadata of each participant in totals or items. 'NA' will be inserted to UserNames which are not in ind_metadata.
```{r}
  items_f_id_s_m <- merge(x=items_f_id_s, y=ind_metadata, by="UserName", all.x=T)
```
  
Check that the items data and metadata are merged.
```{r, eval=FALSE}
  head(items_f_id_s_m)
```

<details>
  <summary>Click to expand output</summary>
```{r, echo=FALSE, eval=TRUE}
head(items_f_id_s_m)
```
</details>

<br>

Save the merged dataframe as a .txt file.
```{r}
  write.table(items_f_id_s_m, "VVKAJ_Items_f_id_s_m.txt", sep="\t", row.names=F, quote=F)
```

<br>

### Generate new totals file if any edits were made to the items file. 

Use one of the input files saved above as an input for calculating totals for. Specify which columns have usernames and Recall.No., which is the number of recorded days. 
```{r}
  GenerateTotals(inputfn = "VVKAJ_Items_f_id_s_m.txt", 
                 User.Name = 'UserName', 
                 Recall.No = 'RecallNo',
                 outfn = "VVKAJ_Tot.txt")
```

Load the total file generated above.
```{r}
  new_totals <- read.table("VVKAJ_Tot.txt", header=T, sep="\t")
```

The number of rows should be {No. of users x No. days}. For the example data, 15 users x 3 days = 45 rows (observations).
```{r}
  nrow(new_totals) 
```

View the new_totals.
```{r, eval=FALSE}
  head(new_totals)
```

<details>
  <summary>Click to expand output</summary>
```{r, echo=FALSE, eval=TRUE}
head(new_totals)
```
</details>

<br>

### <Optional> Add the participants' metadata back to totals.

Load ind_metadata.txt if you have not done so.
```{r}
  ind_metadata <- read.table("ind_metadata.txt", sep="\t", header=T)
```

Add this metadata of each participant to totals. 'NA' will be inserted to UserNames which are not in ind_metadata. 
```{r}
  new_totals_m <- merge(x=new_totals, y=ind_metadata, by="UserName", all.x=T)
```
  
Check that the items data and metadata are merged.
```{r, eval=FALSE}
  head(new_totals_m)
```

<details>
  <summary>Click to expand output</summary>
```{r, echo=FALSE, eval=TRUE}
head(new_totals_m)
```
</details>

<br>

Save the merged dataframe as a .txt file.
```{r}
  write.table(new_totals_m, "VVKAJ_Tot_m.txt", sep="\t", row.names=F, quote=F)
```

<br>

### QC totals data

Look for outliers in your totals.

Note that input dataframe (QCtotals) will be overwritten after outlier removal.
  
Load your totals if necessary - to be used as input for QC.
```{r}
  new_totals <- read.table("VVKAJ_Tot_m.txt", sep="\t", header=T)
```
    
Define your totals dataset to be used as input.
```{r}
  QCtotals <- new_totals  
```

Flag if KCAL is <600 or >5700 --> ask remove or not --> if yes, remove those rows
```{r}
  QCOutliers(input.data = QCtotals, target.colname = "KCAL", min = 600, max = 5700)
```

Flag if PROT is <10 or >240 --> ask remove or not --> if yes, remove those rows
```{r}
  QCOutliers(input.data = QCtotals, target.colname = "PROT", min = 10, max = 240)
```

Flag if TFAT is <15 or >230 --> ask remove or not --> if yes, remove those rows
```{r}
  QCOutliers(input.data = QCtotals, target.colname = "TFAT", min = 15, max = 230)
```

Flag if VC (Vitamin C) is <5 or >400 --> ask remove or not --> if yes, remove those rows
```{r}
  QCOutliers(input.data = QCtotals, target.colname = "VC", min = 5, max = 400)
```

You may find numerous potential outliers here. Then, click "No", and view those outliers with their other nutrient intake information by running the following;
```{r}
  VC_outliers <- subset(new_totals, VC < 5 | VC > 400)    
```

Sort in the order of VC and show only the specified variables.
```{r}
  VC_outliers[order(VC_outliers$BCAR, decreasing = T),
              c('UserName', 'KCAL', 'VC', 'V_TOTAL', 'V_DRKGR', 'F_TOTAL')]  # F is fruits.
```

Save as "Totals_m_QCed.txt"
```{r}
  write.table(QCtotals, "VVKAJ_Tot_m_QCed.txt", sep="\t", quote=F, row.names=F)
```

<br>

---
  
Come back to the main directory before you start running another script.
```{r, eval=FALSE}
  setwd(main_wd)
```
  
