---
title: "Visualize food tree"
output: html_document
---

One can visualize food trees by using the ggtree package.

<br>

Name the path to DietR directory where input files are pulled.
```{r}
  main_wd <- "~/GitHub/DietR"
```

If you have not downloaded and installed the ggtree package yet, you can do so by first installing BiocManager (if you have not done so):
```{r, eval=FALSE}
  if (!require("BiocManager", quietly = TRUE))install.packages("BiocManager")
```
  
Then, use BiocManager to install the "ggtree" package.
```{r, eval=FALSE}
  BiocManager::install("ggtree")
```

Load the necessary functions.
```{r}
  source("lib/specify_data_dir.R")
  source("lib/viz_food_tree.R")
```

You can come back to the main directory by:
```{r, eval=FALSE}
  setwd(main_wd)
```

<br>

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_knit$set(root.dir = 'eg_data/VVKAJ/Foodtree')
```

Go to the "Foodtree" directory where the tree files are saved.
```{r, eval = FALSE}
  SpecifyDataDirectory(directory.name= "eg_data/VVKAJ/Foodtree")
```
<!-- [NOTE] SpecifyDataDirectory function is not executed by R markdown, but just showing  -->
<!--  the code so that this tutorial will have the same code as the R script.  -->

Load the generated food tree. This will load the .nwk file and save it as a tree object called "tree". It is OK to see a message saying: 
Found more than one class "phylo" in cache; using the first, from namespace 'phyloseq' also defined by 'tidytree'.
```{r}
  tree <- read.tree("VVKAJ_Items_f_id_s_m_ff_reduced_4Lv.tree.nwk")

  tree
```

<br> 

Prepare node labels of L1 for plotting. It assumes that the tree file has 9 L1 levels.
```{r}
  PrepFoodTreePlots(input.tree=tree)
```

Create a color-coded and annotated food tree with 9 L1 levels. Choose either 'circular' or 'radial' for layout. It is OK to see some warning messages about Coordinate system and scale for 'y' already being present.
```{r}
  VizFoodTree(input.tree=tree, layout="radial")
```

Look at the color-coded and annotated food tree, saved as annotated_tree.
```{r}
  annotated_tree
```

Save the tree as a PDF file. 
```{r}
ggsave("VVKAJ_Items_f_id_s_m_ff_reduced_4Lv.tree.pdf", annotated_tree, device="pdf", width=6, height=6, units="in", dpi=300)
```

---

<br>

Come back to the main directory.
```{r}
  setwd(main_wd)  
```

