---
title: "Load and clean ASA24 data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_knit$set(root.dir = 'eg_data/VVKAJ')
```

### What this script does:
1. Use Metadata 1 to filter out individuals.
2. Remove users that has only a small number of totals (days of record). - if you know which one to remove.  
3. Look for outliers in your totals by nutrient consumed on each day.

Name the path to DietR directory where input files are pulled.
```{r}
  main_wd <- "~/GitHub/DietR"
  main_wd
```

Load the necessary functions
```{r}
  source("../../lib/specify_data_dir.R")
  source("../../lib/load_clean_ASA24.R")
  source("../../lib/format.file.R")
```

Specify the directory where the data is.
```{r, results = 'hide'}
  SpecifyDataDirectory(directory.name= "eg_data/VVKAJ/")
```

<!-- Define the path to items data in main_wd. -->
<!-- ```{r} -->
<!-- items_path <- paste("/eg_data/VVKAJ/VVKAJ_Items.csv", sep="") -->
<!-- print(items_path) -->
<!-- ``` -->

### Load ASA24 Items data
Load items data which is in a csv format.
```{r}
items_raw <- read.csv("VVKAJ_Items.csv", sep = ",", header=T)
```

Show the first two lines of the items data.
```{r}
head(items_raw, 2)
```

<!-- Save it as a .txt file. -->

<!-- This will be saved in GitHub/DietR_gh_pages/ directory. -->
<!-- ```{r} -->
<!--   write.table(items_raw, "eg_data/VVKAJ/VVKAJ_Items.txt", sep="\t", row.names=F) -->
<!-- ``` -->

<!-- Special characters such as "'", ",", "%" may interfere correct data loading; thus, we replace them with an underscore "_". -->

<!-- This function can only take .txt files as input. Specify column(s) to be processed in the "columns" argument and the output file name in the outfn argument; "_f" stands for "formatted". -->
<!-- ```{r} -->
<!--   format.file(filename = "eg_data/VVKAJ/VVKAJ_Items.txt", -->
<!--               columns  = "eg_data/VVKAJ/Food_Description", -->
<!--               outfn    = "eg_data/VVKAJ/VVKAJ_Items_f.txt") -->
<!-- ``` -->

<!-- Add SampleID with a desired prefix, and save it as a txt file. SampleIDs are IDs unique to each combination of users and day. -->
<!-- ```{r} -->
<!--   AddSampleIDtoItems(input.fn="eg_data/VVKAJ/VVKAJ_Items_f.txt", user.name="UserName", recall.no="RecallNo", -->
<!--                      prefix="vvkaj.", out.fn="eg_data/VVKAJ/VVKAJ_Items_f_id.txt") -->
<!-- ``` -->

<!-- Load the formatted Items file. -->
<!-- ```{r} -->
<!--   items_f_id <- read.table("eg_data/VVKAJ/VVKAJ_Items_f_id.txt", sep="\t", header=T) -->
<!-- ``` -->

<!-- A combination of the specified prefix and sequential number should be added. -->
<!-- ```{r} -->
<!--   head(items_f_id) -->
<!-- ``` -->

<!-- Ensure your items file has the expected dimensions (number of rows x number of columns, shown as number of obs. and number of variables) in the environment window of R Studio, or you can also check the dimension of items_f by using the dim() function. -->
<!-- ```{r} -->
<!--   dim(items_f_id) -->
<!-- ``` -->

<!-- ### <Optional> Use individuals_to_remove.txt to filter out users marked as Remove = yes. -->
<!-- Load your metadata that has information about which UserName(s) to remove. -->
<!-- ```{r} -->
<!-- ind_to_rm_path <- paste(main_wd, "/eg_data/VVKAJ/individuals_to_remove.txt", sep="") -->
<!-- ind_to_rm <- read.table(ind_to_rm_path, sep="\t", header=T) -->
<!-- ``` -->

<!-- Take a look. Metadata for this purpose (ind_to_rm) has UserName and which one to be removed: -->
<!-- ```{r} -->
<!--   ind_to_rm -->
<!-- ``` -->

<!-- Show which has "yes" in the "Remove" column. -->
<!-- ```{r} -->
<!--   subset(ind_to_rm, Remove == "yes") -->

<!-- ``` -->

<!-- Remove the specified individuals. -->
<!-- The output will be saved as a text file with the specified name. -->
<!-- This assumes the usernames are in UserName column, and will print which user(s) will be removed. -->
<!-- ```{r} -->
<!--   RemoveRows(data=items_f_id, metadata.file= ind_to_rm, -->
<!--              output.name= "VVKAJ_Items_f_id_s.txt") -->
<!-- ``` -->

<!-- Load the output for further processing. -->
<!-- ```{r} -->
<!--   items_f_id_s <- read.table("VVKAJ_Items_f_id_s.txt", header=T, sep="\t") -->
<!-- ``` -->

<!-- Show unique usernames in items_f_id_s and confirm "VVKAJ116" has been removed. -->
<!-- ```{r} -->
<!--   unique(items_f_id_s$UserName) -->
<!-- ``` -->


