---
title: "ASA24 No_1 load and clean ASA24 data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

## What this script does:
1. Use Metadata 1 to filter out individuals.
3. Remove users that has only a small number of totals (days of record). - if you know which one to remove.  
2. Look for outliers in your totals by nutrient consumed on each day.

## How to specify path to data.
```{r}
shape_path <- "~/GitHub/DietR"
myfilepath <- paste(shape_path, "/eg_data/VVKAJ/VVKAJ_Items.txt", sep="")
print(myfilepath)
```

## How to load data.
```{r}
myRobject <- read.table(myfilepath, sep="\t", header=T)

head(myRobject)
```

What you need to to: 
Need to hide the part to:
1. Set working directory
2. Load necessary functions and packages. (because the paths aren't correct for users)
3. Load items data etc.

### Load the necessary functions
```{r}
source(paste(shape_path, "/lib/specify_data_dir.R", sep=""))
```

### Load the necessary packages
```{r}
library(ggplot2)
```


need to do something like:
"Load this file in your working directory called "dietary_patterns".
REAL CODE = read.table(" ") 

# Next section
  
```{r}
# for loop solution
# i becomes the number 1 to 100 at each step of the loop


a <- length(100) # make empty variable, set length to 100
for (i in 1:100){
  a[i] <-i #assigns the number in i, to the ith index of a
}

print(a)

# for loop solution #2

a<-c() #create empty variable using combine command
for (i in 1:100){
  a<-c(a,i) # keeps combining a with itself and the new number in i
}
print(a)

# seq solution
a <- seq(1,100,1) # look up help for seq using ?seq() in console
print(a)

```

