---
title: "ASA24 Data Overview"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Name the path to DietR directory where input files are pulled.
```{r}
  main_wd <- "~/GitHub/DietR"
  main_wd
```

Load the necessary functions
```{r}
  source(paste(main_wd, "/lib/specify_data_dir.R", sep=""))
  source(paste(main_wd, "/lib/load_clean_ASA24.R", sep=""))
  source(paste(main_wd, "/lib/format.file.R", sep=""))
```

Define the path to items data in main_wd.
```{r}
items_path <- paste(main_wd, "/testsheet.csv", sep="")
print(items_path)
```

### Load ASA24 Items data
Load items data which is in a csv format.
```{r}
items_raw <- read.csv(items_path, sep = ",", header=T)
```

Show the first two lines of the items data.
```{r}
head(items_raw, 2)
```


### Summarize and visualize ASA24 items and totals data.

### Sample codes
```{r cars}
summary(cars)
```

### Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
