---
title: "Perform ordination"
output: html_document
---

In this section, we will take the phylogeny of food items into account in clustering individuals according to their dietary data. In order to do so, we will use [the phyloseq package](https://joey711.github.io/phyloseq/index.html), which uses phylogeny of microbes and their abundance.  We will replace microbes with food items consumed by our dietary study participants.

Name the path to DietR directory where input files are pulled.
```{r}
  main_wd <- "~/GitHub/DietR"
```

<br>

If you have not downloaded and installed the phyloseq package yet, you can do so by first installing BiocManager (if you have not done so):
```{r, eval=FALSE}
  if (!require("BiocManager", quietly = TRUE)) install.packages("BiocManager")
```

Then download and install the phyloseq package.
```{r, eval=FALSE}
  BiocManager::install("phyloseq")
```

<br>

Load the necessary packages.
```{r}
  library(phyloseq)
  library(ggtree)
```

Load the necessary functions.
```{r}
  source("lib/specify_data_dir.R")
  source("lib/ordination.R")
```

You can come back to the main directory by:
```{r}
  setwd(main_wd)
```

<br>

### Create a phyloseq object for ordination




