---
title: "Visualize ordination"
output: html_document
---

Name the path to DietR directory where input files are pulled.
```{r}
main_wd <- "~/GitHub/DietR"
```

<br>

Load the necessary packages.
```{r}
library(ggplot2)
```

Load the necessary functions.
```{r}
source("lib/specify_data_dir.R")
source("lib/ggplot2themes.R")
```

Load the distinct 100 colors for use.
```{r}
distinct100colors <- readRDS("lib/distinct100colors.rda")
```

You can come back to the main directory by:
```{r}
setwd(main_wd)
```

<br>

### Load ordination results

Change to the folder called "Unifrac" in your "VVKAJ" folder.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_knit$set(root.dir = 'eg_data/VVKAJ/Unifrac')
```

Specify the directory where the data is.
```{r, eval = FALSE}
SpecifyDataDirectory(directory.name= "eg_data/VVKAJ/Unifrac/")
```
<!-- [NOTE] SpecifyDataDirectory function is not executed by R markdown, but just showing  -->
<!--  the code so that this tutorial will have the same code as the R script.  -->
  
Read in the metadata and users' Axis values. 
```{r}
meta_usersdf <- read.table("4Lv_ordinated_Weighted_meta_users.txt", header=T)
```

Take a look at meta_usersdf_loaded.
```{r}
head(meta_usersdf,3)
```

Read in the eigenvalues for axis labels of biplots.
```{r}
eigen_loaded <- read.table("4Lv_ordinated_Weighted_eigen_percent.txt", header=T)
```

---

### Plot Axis 1 and Axis 2 to show the separation of samples colored by UserName, gender, timing, etc. in the metadata

#### By UserName
```{r}
by_user <- ggplot(meta_usersdf, aes(x=Axis.1, y=Axis.2, color=UserName)) +
        geom_point(aes(color=UserName), size=2) + 
        scale_color_manual(values = distinct100colors) + # OR use viridis theme.
        # scale_color_viridis_d() +
        xlab( paste("Axis.1 (", paste(round(eigen_loaded[1,2]*100, 1)), "%)", sep="") ) +
        ylab( paste("Axis.2 (", paste(round(eigen_loaded[2,2]*100, 1)), "%)", sep="") ) +
        no_grid + space_axes + theme(aspect.ratio = 1)
```

```{r out.width="60%"}
by_user
```

Save by_user plot as a PDF.
```{r}
ggsave("4Lv_ordinated_Weighted_Axis12_users.pdf", by_user,
       device="pdf", height=6, width=6, unit="in", dpi=300)
```

Add lines to connect samples in the order in which they appear in the data using `geom_path`.\
**[NOTE]** `geom_line` connects in the order of the variable (small to large) on the x axis, so it could be misleading.
```{r}
by_user_pathconnected <- by_user + geom_path(aes(color = UserName))  
```

```{r out.width="60%"}
by_user_pathconnected
```

Save by_user_pathconnected as a PDF.
```{r}
ggsave("4Lv_ordinated_Weighted_Axis12_users_pathconnected.pdf", 
       by_user_pathconnected, device="pdf", height=6, width=6, unit="in", dpi=300)
```

---

#### By diet (Vegan, Vegetarian, Keto, American, and Japanese)
```{r}
by_diet <- ggplot(meta_usersdf, aes(x=Axis.1, y=Axis.2, color=Diet)) +
  geom_point(aes(color=Diet), size=2) + 
  scale_color_manual(values = distinct100colors) + # OR use viridis theme.
  # scale_color_viridis_d() +
  xlab( paste("Axis.1 (", paste(round(eigen_loaded[1,2]*100, 1)), "%)", sep="") ) +
  ylab( paste("Axis.2 (", paste(round(eigen_loaded[2,2]*100, 1)), "%)", sep="") ) +
  no_grid + space_axes + theme(aspect.ratio = 1)
```

```{r out.width="60%"}
by_diet
```

You can add ellipses at a desired confidence level.
```{r}
by_diet_ellipses <- by_diet + stat_ellipse(level=0.95) 

```

```{r, out.width="60%"}
by_diet_ellipses
```

Save by_user_diet_ellipses as a PDF.
```{r}
ggsave("4Lv_ordinated_Weighted_Axis12_diet_ellipses.pdf", by_diet_ellipses, 
       device="pdf", height=6, width=6, unit="in", dpi=300)
```

Make polygons by diet.
```{r}
by_diet_polygons <- by_diet + geom_polygon(aes(fill = Diet)) + 
  geom_point(aes(color = Diet), size=2) + 
  scale_fill_manual(values = distinct100colors)
```

```{r, out.width="60%"}
by_diet_polygons
```

Save by_diet_polygons as a PDF.
```{r}
ggsave("4Lv_ordinated_Weighted_Axis12_diet_polygons.pdf", 
       by_diet_polygons, device="pdf", height=6, width=6, unit="in", dpi=300)
```

<br>

### Change colors for specific user(s) 

#### [A] Highlight one sample with others being grey

Specify which user to highlight; e.g. VVKAJ101.
```{r}
select_point_1 <- subset(meta_usersdf, UserName=="VVKAJ101") 
```

```{r}
panelA <- by_user + 
  geom_point(size=2, color="grey") +  
  geom_point(data=select_point_1, aes(x=Axis.1, y=Axis.2), color="black", size=2) 
```

```{r, out.width="60%"}
panelA
```

Save the panel as a PDF.
```{r}
ggsave("4Lv_ordinated_Weighted_Axis12_users_VVKAJ101_grey.pdf", 
       panelA, device="pdf", height=6, width=6, unit="in", dpi=300)
```

---

#### [B] Highlight multiple samples with others being grey

Specify colors for each samples to be highlighted in the scale_color_manual argument.
```{r}
select_points <- subset(meta_usersdf, UserName=="VVKAJ101" | UserName=="VVKAJ106" )
```

It is OK to see a message: "Scale for 'colour' is already present.\
Adding another scale for 'colour', which will replace the existing scale."

```{r}
panelB <- 
  by_user + 
  geom_point(data=select_points, aes(x=Axis.1, y=Axis.2, color=as.factor(UserName))) +
  scale_color_manual(values = c("VVKAJ101"="red", "VVKAJ106"="blue")) 
```

```{r, out.width="60%"}
panelB
```

Save the panel as a PDF.
```{r}
ggsave("4Lv_ordinated_Weighted_Axis12_users_VVKAJ101_106_grey.pdf", 
       panelB, device="pdf", height=6, width=6, unit="in", dpi=300)
```

---

#### [C] Highlight one sample; other points will retain their original colors
```{r}
select_point_1 <- subset(meta_usersdf, UserName=="VVKAJ101") 
```

Changing the shape sizes might help find the dots. Note that points may be overlapping
```{r}
panelC <- by_user + 
  geom_point(data=select_point_1, aes(x=Axis.1, y=Axis.2), color="black", size=4) 
```

```{r, out.width="60%"}
panelC
```

Save the panel as a PDF.
```{r}
ggsave("4Lv_ordinated_Weighted_Axis12_users_VVKAJ101_color.pdf", 
       panelC, device="pdf", height=6, width=6, unit="in", dpi=300)
```

---

#### [D] Highlight multiple samples; other points will retain their original colors
```{r}
select_point_1 <- subset(meta_usersdf, UserName=="VVKAJ101") 
select_point_2 <- subset(meta_usersdf, UserName=="VVKAJ106") 

```

```{r}
panelD <- by_user + 
  geom_point(data=select_point_1, aes(x=Axis.1, y=Axis.2), color="black", size=4) +
  geom_point(data=select_point_2, aes(x=Axis.1, y=Axis.2), color="green", size=4) 
```

```{r, out.width="60%"}
panelD
```
  
Save the panel as a PDF.
```{r}
  ggsave("4Lv_ordinated_Weighted_Axis12_users_VVKAJ101_106_color.png", 
         panelD, device="png", height=6, width=6, unit="in", dpi=300)
```

  

