---
title: "Percent kcal of CARB, PROT, TFAT"
output: html_document
---

Name the path to DietR directory where input files are pulled.
```{r}
  main_wd <- "~/GitHub/DietR"
```

Load the necessary functions
```{r}
  source("lib/specify_data_dir.R")
  source("lib/percent_kcal.R")
  source("lib/ggplot2themes.R")
```

Call color palette.
```{r}
  distinct100colors <- readRDS("lib/distinct100colors.rda")
```

You can come back to the main directory by:
```{r, eval=FALSE}
  setwd(main_wd)
```

<br>

## Import data from your data directory

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_knit$set(root.dir = 'eg_data/VVKAJ')
```

Specify the directory where the data is.
```{r, eval = FALSE}
  SpecifyDataDirectory(directory.name= "eg_data/VVKAJ/")
```
<!-- [NOTE] SpecifyDataDirectory function is not executed by R markdown, but just showing  -->
<!--  the code so that this tutorial will have the same code as the R script.  -->

Load the totals data.
```{r}
  totals <- read.table("VVKAJ_Tot_m_QCed.txt",  sep = "\t", header = T)
```

---

Calculate the mean and SD of CARB, PROT, and TFAT.
```{r}
  CPTgramsPerUser(inputfn= totals, user.name = "UserName", recall.no = "RecallNo",
                  outfn="VVKAJ_Tot_m_QCed_CPT_g.txt")
```

The output is not used in visualization below, but you may want to take a look at it.

<br>

Calculate the mean % of energy intake (kcal) and SD of CARB, PROT, and TFAT.
```{r}
  CPTpctKcalPerUser(inputfn=totals, user.name = "UserName", recall.no = "RecallNo", 
                   outfn="VVKAJ_Tot_m_QCed_CPT_kcal.txt")
```

Load the %kcal values 
```{r}
  CPT_kcal <- read.table("VVKAJ_Tot_m_QCed_CPT_kcal.txt", sep="\t", header=T)
```

**[NOTE]** Do not alter the columnnames of CPT_kcal because the plotting functions below assume that CPT_kcal has "UserName", "macronutrient", "n", "mean", and "sd" columns in it.

<br>

---

Plot a barchart without SD. 
```{r}
  # Change the font size if necessary.
  stacked_wo_SD <- StackedwoSD(data= CPT_kcal) + theme(axis.text.x=element_text(size=11))  
```

```{r}
  stacked_wo_SD
```

Save as a .pdf.
```{r}
  ggsave("VVKAJ_Tot_m_QCed_CPT_kcal_wo_SD.pdf", stacked_wo_SD, 
         device="pdf", width=6.2, height=4.2, units="in", dpi=300)
```

---

Plot the "dodge"-type of barchart (3 bars per user, NOT STACKED).
```{r}
  # Change the font size if necessary.
  dodgedtypebarchart <- DodgedBarchart(data= CPT_kcal) + theme(axis.text.x=element_text(size=11))  
```

```{r}
  dodgedtypebarchart
```

Save as a .pdf.
```{r}
  ggsave("VVKAJ_Tot_m_QCed_CPT_kcal_CPT_kcal_dodgedtypebarchart.pdf", dodgedtypebarchart,
         device="pdf", width=9.0, height=4, units="in", dpi=300)
```

---

Using CPT_kcal, create a stacked barchart.
  
Create a vector that contains all the users (individuals).
```{r}
  individuals <- unique(CPT_kcal$UserName)
```

Calculate sd_base and sd_forstacked for stacked barchart. 

Note that this function assumes all users (individuals) have CARB, PROT, and TFAT values.
```{r}
  CalcStackedSD(input.df= CPT_kcal, out.fn= "CPT_kcal_forstacked.txt")
```

Load the saved file that has SD for stacked barchart.
```{r}
  CPT_kcal_forstacked_read <- read.table("CPT_kcal_forstacked.txt", sep="\t", header=T)
```
  
Stacked barchart with SD as error bars.
```{r}
  stacked_with_SD <- StackedWithSD(data=CPT_kcal_forstacked_read) + theme(axis.text.x=element_text(size=11))
```

```{r}
  stacked_with_SD
```

Save as a .pdf.
```{r}
  ggsave("VVKAJ_Tot_m_QCed_CPT_kcal_CPT_kcal_with_SD.pdf", stacked_with_SD, 
         device="pdf", width=6.2, height=4.3, units="in", dpi=300)
```

Change the Y axis scale if necessary. Note that if the error bars of Carbohydrates disappear  after changing the limits of Y axis, it may be because the error bars are higher than the max Y.

Ensure you have enough max Y value to accommodate the error bars.
  
You can also change the breakpoints of the Y axis.
```{r}
  stacked_with_SD + scale_y_continuous(breaks = c(0, 20, 40, 60, 80, 100))
```

<br>

---

Come back to the main directory before you start running another script.
```{r, eval=FALSE}
  setwd(main_wd)
```
  
