---
title: "Percent kcal of CARB, PROT, TFAT"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 3
# code_folding: hide
---

<style type="text/css">
h1.title {
  font-size: 38px;
  font-family: "Tahoma", sans-serif;
  color: DarkSlateGray;
}
h1 { /* Header 1 */
  font-size: 28px;
  color: MidNightBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: MidNightBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  color: MidNightBlue;
}
body{ /* Normal  */
  background-color: rgb(251, 251, 251);
  }
blockquote{ 
  font-size: 14px;
}
</style>

<br>

# Introduction


<br>

# Load functions and packages  

Name the path to DietR directory where input files are pulled.
```{r}
main_wd <- "~/GitHub/DietR"
```

Load the necessary functions.
```{r}
source("lib/specify_data_dir.R")
source("lib/percent_kcal.R")
source("lib/ggplot2themes.R")
```

Call color palette.
```{r}
distinct100colors <- readRDS("lib/distinct100colors.rda")
```

You can come back to the main directory by:
```{r, eval=FALSE}
setwd(main_wd)
```

Specify the directory where the data is.
```{r, eval = FALSE}
SpecifyDataDirectory(directory.name= "eg_data/VVKAJ/")
```

<!-- This is the R-markdown code to change directories. -->
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_knit$set(root.dir = 'eg_data/VVKAJ')
```

<br>





<br>  

---
Come back to the main directory before you start running another script.
```{r, eval=FALSE}
  setwd(main_wd)
```

<!-- This is to remove big white space at the end of the rendered html, which is produced due to toc_floating.  Source: https://stackoverflow.com/questions/52933437/how-to-remove-white-space-at-the-end-of-an-rmarkdown-html-output -->
<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>
