---
title: "Data overview"
output: html_document
---

Name the path to DietR directory where input files are pulled.
```{r}
  main_wd <- "~/GitHub/DietR"
```

Load the necessary functions
```{r}
  source("lib/specify_data_dir.R")
  source("lib/data_overview.R")
  source("lib/ggplot2themes.R")
```

Call color palette.
```{r}
  distinct100colors <- readRDS("lib/distinct100colors.rda")
```

You can come back to the main directory by:
```{r, eval=FALSE}
  setwd(main_wd)
```

<br>

### Load and analyze (QC-ed) ASA24 items data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_knit$set(root.dir = 'eg_data/VVKAJ')
```

Specify the directory where the data is.
```{r, eval = FALSE}
  SpecifyDataDirectory(directory.name= "eg_data/VVKAJ/")
```
<!-- [NOTE] SpecifyDataDirectory function is not executed by R markdown, but just showing  -->
<!--  the code so that this tutorial will have the same code as the R script.  -->

Load your items data to be analyzed.
"_f_id_s_m" stands for: "food names formatted", "SampleID added", "selected individuals", and "metadata merged".
```{r}
  items_f_id_s_m <- read.table("VVKAJ_Items_f_id_s_m.txt", sep="\t", header=T)
```

```{r}
  head(items_f_id_s_m)
```

<br>

### Generate sum of foods consumed by each user, by day, and by occasion


<br>

---
  
Come back to the main directory before you start running another script.
```{r, eval=FALSE}
  setwd(main_wd)
```
  



