---
title: "Title"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 3
# code_folding: hide
---

<style type="text/css">
h1.title {
  font-size: 38px;
  font-family: "Tahoma", sans-serif;
  color: DarkSlateGray;
}
h1 { /* Header 1 */
  font-size: 28px;
  color: MidNightBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: MidNightBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  color: MidNightBlue;
}
body{ /* Normal  */
  background-color: rgb(251, 251, 251);
  }
blockquote{ 
  font-size: 14px;
}
</style>

<br>

# Introduction


<br>

# Load functions and packages  

Name the path to DietR directory where input files are pulled.
```{r}
main_wd <- "~/GitHub/DietR"
```

Install the SASxport package if it is not installed yet.
```{r, eval=FALSE}
if (!require("SASxport", quietly = TRUE)) install.packages("SASxport")
```

Load necessary packages.
```{r}
library(SASxport)
```

Load the necessary functions.
```{r}
source("lib/specify_data_dir.R")
```

Load the distinct 100 colors for use.
```{r}
distinct100colors <- readRDS("lib/distinct100colors.rda")
```

You can come back to the main directory by:
```{r, eval=FALSE}
setwd(main_wd)
```

Specify the directory where the data is.
```{r, eval=FALSE}
SpecifyDataDirectory(directory.name = "eg_data/NHANES")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_knit$set(root.dir = 'eg_data/NHANES')
```

<br>





<br>  

---

Come back to the main directory.
```{r, eval=FALSE}
setwd(main_wd)
```

<!-- This is to remove big white space at the end of the rendered html, which is produced due to toc_floating.  Source: https://stackoverflow.com/questions/52933437/how-to-remove-white-space-at-the-end-of-an-rmarkdown-html-output -->
<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>
