---
title: "Load and clean NHANES data"
output: html_document
---

NHANES is a national survey conducted with some thousands of participants across the US over two days. NHANES data is published every other year and consists of various kinds of data including dietary intake, body measurements, laboratory-taken data. Here, we will explore the dietary intake (food items) of participants and look at the relationship between dietary patterns and diabetic status in [NHANES 2015-2016](https://wwwn.cdc.gov/nchs/nhanes/continuousnhanes/default.aspx?BeginYear=2015).

Refer to [the NHANES data and documentation](https://wwwn.cdc.gov/nchs/nhanes/Default.aspx) for more information about each release (version) of NHANES and their variables measured.

Data are in .XPT format, and you will need a specific package, [SASxport](https://github.com/r-gregmisc/SASxport), to import it to R. The example data we will use here were downloaded from the NHANES 2015-2016 page and are in the “Raw_data” folder in the “NHANES” folder.

Open “02_load_clean_NHANES_food.R” script in the “users” folder and load the packages needed as per the directions.

<br>

Name the path to DietR directory where input files are pulled.
```{r}
  main_wd <- "~/GitHub/DietR"
```

Load the necessary functions.
```{r}
  source("lib/specify_data_dir.R")
  source("lib/load_clean_NHANES.R")
  source("lib/prep_data_for_clustering.R")
  source("lib/Food_tree_scripts/format.foods.r")
```

Install the SASxport package if it is not installed yet.
```{r, eval=FALSE}
  if (!require("SASxport", quietly = TRUE)) install.packages("SASxport")
```

Load the necessary packages.
```{r}
library(SASxport)
```

You can come back to the main directory by:
```{r, eval=FALSE}
  setwd(main_wd)
```

---

### Prepare food codes, FPED, and food items files

#### Load and prepare food codes

Download food code data from NHANES website and save it in "Raw_data" folder.
```{r, eval=FALSE}
download.file("https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/DRXFCD_I.XPT", 
              destfile= "eg_data/NHANES/Raw_data/FoodCodes_DRXFCD_I.XPT", mode="wb")
```

Prepare the food code table - replace special characters with "_" or "and".

Format the food table and save it as a .txt file.
```{r}
PrepareFoodCodeTable(raw.food.code.table = "eg_data/NHANES/Raw_data/FoodCodes_DRXFCD_I.XPT", 
                     out.fn = "eg_data/NHANES/FoodCodes_DRXFCD_I_f.txt")  

```

Load the formatted foodcode table.
```{r}
foodcodetable_f <- read.table("eg_data/NHANES/FoodCodes_DRXFCD_I_f.txt", sep="\t", header=T)
```

Show the first 10 rows of the output and ensure special characters are gone; e.g., "MILK, REDUCED FAT (2%)" is now "MILK, REDUCED FAT (2_)".
```{r}
foodcodetable_f[1:10, ]
```

---

#### Load and prepare FPED

FPED has the composition of nutrients and food categories of each food item coded by food codes. FPED can be downloaded from [USDA-ARS FPED databases](https://www.ars.usda.gov/northeast-area/beltsville-md-bhnrc/beltsville-human-nutrition-research-center/food-surveys-research-group/docs/fped-databases/).

You need to use the correct version of FPED. With NHANES, use FPED with the same release year as the year of NHANES you are analyzing. For this tutorial, FPED was downloaded from NHANES the FPED table is formatted by renaming the variables with R-loadable ones (e.g., "F_CITMLB (cup eq.)" --> "F_CITMLB" and saved as "FPED_1516_forR.txt".

Load FPED15-16, needed for the AddFoodCat function.
```{r}
FPED <- read.table("eg_data/NHANES/FPED/FPED_1516_forR.txt", sep="\t", header=T)
```

Check the first two rows of FPED.
```{r}
head(FPED, 2)
```

**Important!** Change the food code column name as Food_code.
```{r}
colnames(FPED)[1] <- "Food_code"
```

---

### Load "food items" data and add food descriptions

#### Download food items data

Food data can be downloaded from the NHANES website.\
Name the file and destination. `mod="wb"` is needed for the Windows OS; other OS users may need to delete it.

Download day 1 data.
```{r, eval=FALSE}
download.file("https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/DR1IFF_I.XPT",
              destfile= "eg_data/NHANES/DR1IFF_I.XPT", mode="wb")
```

Download day 2 data.
```{r, eval=FALSE}
download.file("https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/DR2IFF_I.XPT",
              destfile= "eg_data/NHANES/DR2IFF_I.XPT", mode="wb")
```

**[NOTE]** Different alphabets are used on the variables' names in different release of NHANES data.  Therefore, you will need to change the alphabet (and potentially the other parts of the variable names) in order to run this script with other releases of NHANES. For example, DR1IFDCD is a column name for the food code used in the NHANES 2015-2016, and the alphabet for this release is "I".

#### Load and prepare Day 1 food items data

Import items data Day 1, add food item descriptions, and save it as a txt file. *** OUTPUT WILL LIKELY BE A HUGE FILE. ***
```{r}
ImportNHANESFoodItems(data.name="eg_data/NHANES/DR1IFF_I.XPT",
                      food.code.column = "DR1IFDCD",
                      food.code.table = foodcodetable_f,
                      out.fn = "eg_data/NHANES/DR1IFF_I_d.txt") # 'd' stands for food descriptions
```

Load the saved food items file.
```{r}
Food_D1 <- read.table("eg_data/NHANES/DR1IFF_I_d.txt", sep="\t", header=T)
```

Count the number of participants - should be 8,505 people.
```{r}
length(unique(Food_D1$SEQN))
```

Add the food category info and serving for each item. *** THIS WILL TAKE A FEW MOMENTS. ***\
It is OK to see a message saying "NAs introduced by coercion". NAs will be removed later in the filtering process.

```{r}
AddFoodCat(input.food= Food_D1,
           fped= FPED,
           grams= "DR1IGRMS",
           out.fn= "eg_data/NHANES/Food_D1_FC.txt")
```

---

#### Load and prepare Day 2 food items data

Import items data Day 2, add food item descriptions, and save it as a txt file.
```{r}
ImportNHANESFoodItems(data.name="eg_data/NHANES/DR2IFF_I.XPT",
                      food.code.column = "DR2IFDCD",
                      food.code.table = foodcodetable_f,
                      out.fn = "eg_data/NHANES/DR2IFF_I_d.txt")
```

Add food item description and save it as a txt file.
```{r}
Food_D2 <- read.table("eg_data/NHANES/DR2IFF_I_d.txt", sep="\t", header=T)
```

Count the number of participants - should be 7027 people.
```{r}
length(unique(Food_D2$SEQN))
```

Do the same for Day 2. Add the food items info and serving for each item. *** THIS WILL TAKE A FEW MOMENTS. ***\
It is OK to see a message saying "NAs introduced by coercion." NAs will be removed later in the filtering process.
```{r}
AddFoodCat(input.food= Food_D2,
           fped= FPED,
           grams= "DR2IGRMS",
           out.fn= "eg_data/NHANES/Food_D2_FC.txt")
```

<br>

### Load the Food_Dx_FC which has the food category data

#### Change column names to more intuitive ones

Load Food Day 1 with Food Category *** THIS WILL BE A HUGE TABLE. ***
```{r}
Food_D1_FC <- read.table("eg_data/NHANES/Food_D1_FC.txt", sep="\t", header=T)
```

Load Food Day 2 with Food Category *** THIS WILL BE A HUGE TABLE. ***
```{r}
Food_D2_FC <- read.table("eg_data/NHANES/Food_D2_FC.txt", sep="\t", header=T)
```

Change the colnames for downstream analyses.
```{r}
names(Food_D1_FC)[names(Food_D1_FC) == "DR1IFDCD"] <- "FoodCode"
names(Food_D1_FC)[names(Food_D1_FC) == "DR1IGRMS"] <- "FoodAmt"
names(Food_D1_FC)[names(Food_D1_FC) == "DRXFCLD"] <- "Main.food.description"

names(Food_D2_FC)[names(Food_D2_FC) == "DR2IFDCD"] <- "FoodCode"
names(Food_D2_FC)[names(Food_D2_FC) == "DR2IGRMS"] <- "FoodAmt"
names(Food_D2_FC)[names(Food_D2_FC) == "DRXFCLD"] <- "Main.food.description"
```

Ensure the column names are changed.
```{r}
colnames(Food_D1_FC)
```

Save after changing the column names. "cc" stands for column names changed.
```{r}
  write.table(Food_D1_FC, "eg_data/NHANES/Food_D1_FC_cc.txt", sep="\t", row.names=F, quote=F)
  write.table(Food_D2_FC, "eg_data/NHANES/Food_D2_FC_cc.txt", sep="\t", row.names=F, quote=F)
```

#### Replace special characters with "_" using the FormatFoods function

The FotmatFoods function adds Main.Food.Description where special characters are removed/replaced, the previous Main.Food.Description as Old.Main.Food.Description, ModCode, and FoodID. $FoodID is a character vector, but has .0 at the end.

MAKE SURE YOU HAVE `dedupe=F`. If true (default!), duplicated foods will be removed!
```{r}
FormatFoods(input_fn="eg_data/NHANES/Food_D1_FC_cc.txt", output_fn= "eg_data/NHANES/Food_D1_FC_cc_f.txt", dedupe=F)

FormatFoods(input_fn="eg_data/NHANES/Food_D2_FC_cc.txt", output_fn= "eg_data/NHANES/Food_D2_FC_cc_f.txt", dedupe=F)
```

